---
title: "Swiss public transport stops"
author: "Olivier von Dach"
date: "18/06/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Synopsis

The basic geo-data set for public transport stops comprises public transport stops in Switzerland and additional selected geo-referenced public transport locations that are of operational or structural importance (operating points).

The data can be downloaded [here](https://opendata.swiss/en/dataset/haltestellen-des-offentlichen-verkehrs)

## Data processing

### Loading

```{r}
data <- read.csv("PointExploitation.csv")
```

### Short analysis

```{r}
dim(data)
names(data)
table(data$MoyenTransport)
```

### Coordinate transformation

The locations of the stops are provided using the CH1903 Swiss coordinate system. A mathematical transformation has to be applied since GPS (WGS84) coordinates are required for the plotting.  

The source code for the transformation in R language is [here](https://github.com/ValentinMinder/Swisstopo-WGS84-LV03/blob/master/scripts/r/WGS84_CH1903.R)

```{r}
source("WGS84_CH1903.R")
```

### Cleaning

```{r, echo=FALSE}
library(dplyr)
```

Cleanup of column names, selection of meaningful columns and conversion of coordinates.

```{r}
df <- data %>% 
    select(Nom, MoyenTransport, y_Coord_Est, x_Coord_Nord) %>% 
    mutate(Stop=Nom, Kind=MoyenTransport,
           Latitude=CH.to.WGS.lat(y_Coord_Est, x_Coord_Nord), 
           Longitude=CH.to.WGS.lng(y_Coord_Est, x_Coord_Nord)) %>% 
    select(-MoyenTransport, -Nom, -y_Coord_Est, -x_Coord_Nord)
head(df)
```

## Data plotting

```{r, echo=FALSE}
library(leaflet)
```

### Funicular stops

```{r}
df %>% 
    select(-Stop) %>% 
    filter(Kind=='Funiculaire') %>%
    leaflet() %>%
    addTiles() %>%
    addCircleMarkers(clusterOptions = markerClusterOptions())
```

### Cableway stops

```{r}
df %>% 
     select(-Stop) %>% 
     filter(Kind=='Telepherique') %>%
     leaflet() %>%
     addTiles() %>%
     addCircleMarkers(clusterOptions = markerClusterOptions())
```

